---
title: "Devil's Lake Analysis"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(raster)
library(robustbase)
library(robust)
library(ggplot2)
```

## Time Series Analysis of Devils Lake Data at Minnewaukan

Load the set of predictions for Minnewaukan four mile square generated using a Random Forest classifier with training on the set of images in "devils_lake_training_image.tif". The data have been imputed with a biannual mean estimation rather than using the entire time range for estimation of the mean.

```{r}
df <- read.csv("devils_lake_predictions_lambda0.3.csv")
df <- df %>% mutate(dates = lubridate::ymd(dates)) %>% mutate(jrc = ifelse(dates > lubridate::ymd("2015-12-31"), NA, jrc))
```

Create a simple time series of the number of pixels classified as water over time. 

```{r}
ggplot(df) + 
  geom_point(aes(x = dates, y = pred*9/10), size = 0.5) + 
  ylab("Water extent (1000 m^2)") +
  xlab("Date") +
  scale_x_date(limits = c(lubridate::ymd("1984-01-01"),lubridate::ymd("2019-12-31"))) +
  theme(text = element_text(size=15))
```

Compare timeseries from JRC and SWIM methods

```{r}
require(gridExtra)
theme_update(plot.title = element_text(hjust = 0.5))
plot1 <- ggplot(df) + 
  ggtitle("JRC") +   
  geom_point(aes(x = dates, y = jrc*9/10), size = 0.5) + 
  scale_y_continuous(limits = c(0,max(df$pred*9/10))) +
  ylab(bquote('Water extent (1000 '*m^2*')')) +
  xlab(NULL) +
  scale_x_date(limits = c(lubridate::ymd("1984-01-01"),lubridate::ymd("2019-12-31"))) +
  theme(text = element_text(size=15))
plot2 <- ggplot(df) + 
  ggtitle("CSSM λ = 0.3") +
  geom_point(aes(x = dates, y = pred*9/10), size = 0.5) + 
  scale_y_continuous(limits = c(0,max(df$pred*9/10)), labels=NULL) +
  ylab(NULL) +
  xlab(NULL) +
  scale_x_date(limits = c(lubridate::ymd("1984-01-01"),lubridate::ymd("2019-12-31"))) +
  theme(text = element_text(size=15))
grid.arrange(plot1, plot2, ncol=2)
timeseries <- arrangeGrob(plot1, plot2, ncol=2)
```

Create a complex graphic with different colors for percent image imputed and a trend connecting only those points with < 1% of the image imputed.

```{r}
rcolors = c('#C8102E', colorRampPalette(c("pink", "blue"))(6))

plot(x = (df %>% filter(cloud < 101))$dates, y = (df %>% filter(cloud < 101))$pred, "p", xlab = "Date", ylab = "Total pixels classified as water", pch = 19, col = rcolors[1], main="Total pixels classified as water after imputation")
points(x = (df %>% filter(cloud < 100))$dates, y = (df %>% filter(cloud < 100))$pred, pch = 19, col = rcolors[2])
points(x = (df %>% filter(cloud < 75))$dates, y = (df %>% filter(cloud < 75))$pred, pch = 19, col = rcolors[3])
points(x = (df %>% filter(cloud < 50))$dates, y = (df %>% filter(cloud < 50))$pred, pch = 19, col = rcolors[4])
points(x = (df %>% filter(cloud < 25))$dates, y = (df %>% filter(cloud < 25))$pred, pch = 19, col = rcolors[5])
points(x = (df %>% filter(cloud < 10))$dates, y = (df %>% filter(cloud < 10))$pred, pch = 19, col = rcolors[6])
points(x = (df %>% filter(cloud < 1))$dates, y = (df %>% filter(cloud < 1))$pred, pch = 19, col = rcolors[7])
lines(x = (df %>% filter(cloud < 1))$dates, y = (df %>% filter(cloud < 1))$pred, "l")
legend(col = rcolors, legend = c("100", "[75, 100)", "[50, 75)", "[25, 50)", "[10, 25)", "[1, 10)", "[0, 1)"), "bottomright", pch = 19, title = "Percent imputed", cex = 1, ncol=1)
```

```{r}
hist(x = df$cloud, xlab = "Percent Messing", main = "Distribution of percentage of image missing")
```

## Spatial Analysis of Devils Lake at Minnewaukan

Load images of four mile square at Minnewaukan

```{r}
aniBrick_dl <- brick("../classification/devils_lake_raster_lambda0.3.tif")
aniBrick_dl <- readAll(aniBrick_dl)
```

Determine which image indexes belong to 1984

```{r}
yearly <- brick()

par(mar = c(0.5,0,1,0), mfrow = c(5,7), oma = c(5,0,0,0))
for(i in 1984:2018){
    images_year <- (df %>% filter(year == i))$row
    par(mar = c(0.5,0,1,0))
    plot(sum(aniBrick_dl[[images_year]])/length(images_year), legend = F, axes = F, box = F, main = i, 
         breaks = c(0,0.0001,0.2,0.4,0.6,0.8,1),
         col = c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5)))
    yearly <- addLayer(yearly, sum(aniBrick_dl[[images_year]])/length(images_year))
}

par(fig = c(0, 1, 0, 1), oma = c(1, 0, 0, 0), mar = c(0, 0, 0, 0), new = TRUE)
plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
legend("bottom", ncol=6, legend=c("Always Land",NA,NA,NA,NA,"Always Water"), xpd = TRUE, horiz = TRUE, inset = c(0, 
    0), bty="n", fill=c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5)), title="Classification Frequency", cex=1.25)
```

Animate surface water extent over all years

```{r}
par(mfrow = c(1,1), mar = c(1,1,1,1))
animate(yearly, main = c(1984:2018), n = 1, pause = 1,
        breaks = c(0,0.0001,0.2,0.4,0.6,0.8,1),
        lab.breaks = c("Always\nland",NA,NA,NA,NA,NA,"Always\nwater"),
        col = c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5)))
```

Recognize anomalous images via high spatial correlation between images with similar number of pixels classified as water, look for high correlation in "extreme weather event" images, if spatial pattern is different it is likely one of those images is anomalous.

```{r}
brick <- aniBrick_dl

yearly <- brick()

for(i in 1984:2018){
    images_year <- (df %>% filter(year == i))$row
    yearly <- addLayer(yearly, sum(brick[[images_year]]))
}

fluctuation_water <- c()
fluctuation_land <- c()
c <- 0
dates <- c()
row <- c()
mean_water <- c()
image_water <- c()
water_brick_dl <- brick()
land_brick_dl <- brick()

for(i in 1984:2018){
  c <- c + 1
  images_year <- (df %>% filter(year == i))$row
  print(i)
  mean <- sum(brick[[images_year]])/length(images_year)
  mean <- mean >= 0.5
  for(j in images_year){
    
    stat_1 <- (brick[[j]] - mean)/(yearly[[c]])
    water <- sum(getValues(stat_1 * (stat_1 > 0)), na.rm = T)
    
    stat_2 <- (1 - brick[[j]] - 1 + mean)/(length(images_year)-yearly[[c]])
    land <- sum(getValues(stat_2 * (stat_2 > 0)), na.rm = T)
    
    water_brick_dl <- addLayer(water_brick_dl, stat_1 * (stat_1 > 0))
    land_brick_dl <- addLayer(land_brick_dl, stat_2 * (stat_2 > 0))
    
    mean_water <- c(mean_water, sum(getValues(mean), na.rm=T))
    image_water <- c(image_water, sum(getValues((brick[[j]])), na.rm=T))
    
    fluctuation_water <- c(fluctuation_water, water)
    fluctuation_land <- c(fluctuation_land, land)
    
    dates <- c(dates, as.character(df[df$row == j,]$dates))
    row <- c(row, j)
  }
}

fluctuation_water_dl <- fluctuation_water
fluctuation_land_dl <- fluctuation_land
mean_water_dl <- mean_water
image_water_dl <- image_water
row_dl <- row
dates_dl <- dates

write.csv(cbind(row_dl, fluctuation_water_dl, fluctuation_land_dl, mean_water_dl, image_water_dl, dates_dl), file = "outlier_analysis_dl.csv")

```

```{r}
t <- read.csv("outlier_analysis_dl.csv")
brick <- aniBrick_dl
fluctuation_water <- t$fluctuation_water_dl
fluctuation_land <- t$fluctuation_land_dl
mean_water <- t$mean_water_dl
image_water <- t$image_water_dl
row <- t$row_dl
dates <- t$dates_dl

######################
resp_water <- fluctuation_water[fluctuation_water>0]
water_mod <- glmrob(resp_water ~ image_water[fluctuation_water>0] +
                        poly(lubridate::yday(dates[fluctuation_water>0]),3),
                    family = Gamma(link="log"))

sum(residuals(water_mod, "deviance")^2)
######################

res <- residuals(water_mod, "deviance") #look only at positive residuals indicating very high water fluctuation
q <- 2

plot(df$pred, x = df$dates, pch = 19)

pal <- colorRampPalette(c("red", "pink"))

points(x = df[(df$row %in% row[fluctuation_water>0][which(res > q)]),]$dates[order(resp_water[which(res > q)], (row[fluctuation_water>0])[which(res > q)], decreasing = T)], y = df[(df$row %in% row[fluctuation_water>0][which(res > q)]),]$pred[order(resp_water[which(res > q)], (row[fluctuation_water>0])[which(res > q)], decreasing = T)], pch = 19, col = pal(length(row[fluctuation_water>0][which(res > q)])))

mean_high_outliers <- df[(df$row %in% row[fluctuation_water>0][which(res > q)]),]$row

mean_land <- ncell(brick[[1]]) - mean_water

###################
resp_land <- fluctuation_land[fluctuation_land>0]
land_mod <- glmrob(resp_land ~ image_water[fluctuation_land>0] +
                       poly(lubridate::yday(dates[fluctuation_land>0]),2),
                   family = Gamma(link="log"))

sum(residuals(land_mod, "deviance")^2)
###################

res <- residuals(land_mod) #look only at positive residuals indicating very high land fluctuation
q2 <- 2

pal <- colorRampPalette(c("blue", "lightblue"))

points(x = df[(df$row %in% row[fluctuation_land>0][which(res > q2)]),]$dates[order(resp_land[which(res > q2)], (row[fluctuation_land>0])[which(res > q2)], decreasing = T)], y = df[(df$row %in% row[fluctuation_land>0][which(res > q2)]),]$pred[order(resp_land[which(res > q2)], (row[fluctuation_land>0])[which(res > q2)], decreasing = T)], pch = 19, col = pal(length(row[fluctuation_land>0][which(res > q2)])))


mean_low_outliers <- df[(df$row %in% row[fluctuation_land>0][which(res > q2)]),]$row

mean_outliers <- c(mean_high_outliers, mean_low_outliers)

which(df[,'row'] %in% mean_high_outliers)
which(df[,'row'] %in% mean_low_outliers)
```

```{r}
require(gridExtra)
theme_update(plot.title = element_text(hjust = 0.5))
plot1 <- ggplot(df) + 
  ggtitle("JRC") +   
  geom_point(aes(x = dates, y = jrc*9/10), size = 0.25) + 
  scale_y_continuous(limits = c(0,max(df$pred*9/10))) +
  ylab(bquote('Water extent (1000 '*m^2*')')) +
  xlab(NULL) +
  scale_x_date(limits = c(lubridate::ymd("1984-01-01"),lubridate::ymd("2019-12-31"))) +
  theme(text = element_text(size=15))
plot2 <- ggplot(df %>% filter(!(row %in% mean_outliers))) + 
  ggtitle(bquote('SWIM '~lambda~'= 0.3')) + 
  geom_point(aes(x = dates, y = pred*9/10), size = 0.25) + 
  scale_y_continuous(limits = c(0,max(df$pred*9/10)), labels=NULL) +
  ylab(NULL) +
  xlab(NULL) +
  scale_x_date(limits = c(lubridate::ymd("1984-01-01"),lubridate::ymd("2019-12-31"))) +
  theme(text = element_text(size=15))
grid.arrange(plot1, plot2, ncol=2)
timeseries2 <- arrangeGrob(plot1, plot2, ncol=2, widths = c(1.2,1))
ggsave(timeseries2, filename = "devils_lake_timeseries_outliers_rm.eps", dpi = 1200, width = 7, height = 3, units = "in")
```

```{r}
require(gridExtra)
theme_update(plot.title = element_text(hjust = 0.5))
plot1 <- ggplot(df) + 
  ggtitle("JRC") +   
  geom_point(aes(x = dates, y = jrc*9/10), size = 0.5) + 
  scale_y_continuous(limits = c(0,max(df$pred*9/10))) +
  ylab(bquote('Water extent (1000 '*m^2*')')) +
  xlab(NULL) +
  scale_x_date(limits = c(lubridate::ymd("1984-01-01"),lubridate::ymd("2019-12-31"))) +
  theme(text = element_text(size=15))
plot2 <- ggplot(df) + 
  ggtitle(bquote(list("SWIM", lambda==0.3))) +
  geom_point(aes(x = dates, y = pred*9/10), size = 0.5) + 
  scale_y_continuous(limits = c(0,max(df$pred*9/10)), labels=NULL) +
  ylab(NULL) +
  xlab(NULL) +
  scale_x_date(limits = c(lubridate::ymd("1984-01-01"),lubridate::ymd("2019-12-31"))) +
  theme(text = element_text(size=15))
plot3 <- ggplot(df %>% filter(!(row %in% mean_outliers))) + 
  ggtitle("SWIM identify outliers") +
  geom_point(aes(x = dates, y = pred*9/10), size = 0.5) + 
  geom_point(data = df %>% filter((row %in% mean_outliers)), aes(x = dates, y = pred*9/10), colour="red", size=0.5, shape=4) +
  scale_y_continuous(limits = c(0,max(df$pred*9/10))) +
  ylab(bquote('Water extent (1000 '*m^2*')')) +
  xlab(NULL) +
  scale_x_date(limits = c(lubridate::ymd("1984-01-01"),lubridate::ymd("2019-12-31"))) +
  theme(text = element_text(size=15))
plot4 <- ggplot(df %>% filter(!(row %in% mean_outliers))) + 
  ggtitle("SWIM outliers removed") +
  geom_point(aes(x = dates, y = pred*9/10), size = 0.5) + 
  scale_y_continuous(limits = c(0,max(df$pred*9/10)), labels=NULL) +
  ylab(NULL) +
  xlab(NULL) +
  scale_x_date(limits = c(lubridate::ymd("1984-01-01"),lubridate::ymd("2019-12-31"))) +
  theme(text = element_text(size=15))
grid.arrange(plot1, plot2, plot3, plot4, ncol=2, widths = c(1.2,1))
timeseries3 <- arrangeGrob(plot1, plot2, plot3, plot4, ncol=2, widths = c(1.2,1))
#ggsave(timeseries3, filename = "devils_lake_timeseries_all.eps", dpi = 1200, width = 8.5, height = 8.5, units = "in")
```

```{r}
png(filename = "devils_lake_spatial_outliers_rm.png", width = 8.5, height = 8.5, units = "in", res = 1200)

yearly <- brick()

par(mar = c(0.5,0,1,0), mfrow = c(5,7), oma = c(5,0,0,0))
for(i in 1984:2018){
    images_year <- (df %>% filter(year == i) %>% filter(!(row %in% mean_outliers)))$row
    par(mar = c(0.5,0,1,0))
    plot(sum(aniBrick_dl[[images_year]])/length(images_year), legend = F, axes = F, box = F, main = i, 
         breaks = c(0,0.0001,0.2,0.4,0.6,0.8,1),
         col = c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5)))
    yearly <- addLayer(yearly, sum(brick[[images_year]])/length(images_year))
}

par(fig = c(0, 1, 0, 1), oma = c(1, 0, 0, 0), mar = c(0, 0, 0, 0), new = TRUE)
plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
legend("bottom", ncol=6, legend=c("0%","20%","40%","60%","80%","100%"), xpd = TRUE, horiz = TRUE, inset = c(0, 
    0), bty="n", fill=c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5)), title="Water Frequency", cex=1.25)

dev.off()
```

```{r}

png(filename = "devils_lake_imputation_figure.png", width = 5, height = 9, units = "in", res = 1200)

par(mar = c(0,0,0,0), mfrow = c(3,2), oma = c(10,5,0,0))

y <- 2001; m <- 5; plot(brick('~/GEE/data/devils_lake/devils_lake_final_extras_rm_jrc.tif')[[(df %>% filter(year == y) %>% filter(month == m) %>% filter(!row %in% mean_outliers))$row[1]]], col = c("white", rgb(92,157,91,maxColorValue=255), "blue"), legend = F, axes = F, box = F)
title("JRC", cex.main = 2, line = -2)
mtext("May 2001", side = 2, cex = 1.5, line = 1)

plot(sum(brick[[(df %>% filter(year == y) %>% filter(month == m) %>% filter(!row %in% mean_outliers))$row]], na.rm=T)/length((df %>% filter(year == y) %>% filter(month == m) %>% filter(!row %in% mean_outliers))$row), col = c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5)), legend = F, axes = F, box = F);
title("NRI", cex.main = 2, line = -2)

y <- 2012; m <- 7; plot(brick('~/GEE/data/devils_lake/devils_lake_final_extras_rm_jrc.tif')[[(df %>% filter(year == y) %>% filter(month == m) %>% filter(!row %in% mean_outliers))$row[1]]], col = c("white", rgb(92,157,91,maxColorValue=255), "blue"), legend = F, axes = F, box = F)
mtext("July 2012", side = 2, cex = 1.5, line = 1)

plot(sum(brick[[(df %>% filter(year == y) %>% filter(month == m) %>% filter(!row %in% mean_outliers))$row]], na.rm=T)/length((df %>% filter(year == y) %>% filter(month == m) %>% filter(!row %in% mean_outliers))$row), col = c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5)), legend = F, axes = F, box = F);

y <- 2015; m <- 10; plot(brick('~/GEE/data/devils_lake/devils_lake_final_extras_rm_jrc.tif')[[(df %>% filter(year == y) %>% filter(month == m) %>% filter(!row %in% mean_outliers))$row[1]]], col = c("white", rgb(92,157,91,maxColorValue=255), "blue"), legend = F, axes = F, box = F)
mtext("October 2015", side = 2, cex = 1.5, line = 1)

plot(sum(brick[[(df %>% filter(year == y) %>% filter(month == m) %>% filter(!row %in% mean_outliers))$row]], na.rm=T)/length((df %>% filter(year == y) %>% filter(month == m) %>% filter(!row %in% mean_outliers))$row), col = c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5)), legend = F, axes = F, box = F);

par(fig = c(0, 1, 0, 1), oma = c(1, 0, 0, 0), mar = c(0, 0, 0, 0), new = TRUE)
plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")

legend("bottom", ncol=3, legend=c("0%","20%","40%","60%","80%","100%")[c(1,4,2,5,3,6)], xpd = TRUE, inset = c(0, 
    0), bty="n", fill=c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5))[c(1,4,2,5,3,6)], title="Water Frequency", cex=2)

dev.off()
```

```{r}
yearly <- brick()

par(mar = c(0.5,0,1,0), mfrow = c(3,2), oma = c(5,0,0,0))
for(i in c(1985,1991,1997,2003,2009,2015)){
    images_year <- (df %>% filter(year == i) %>% filter(!(row %in% mean_outliers)))$row
    par(mar = c(0.5,0,1,0))
    plot(sum(aniBrick_dl[[images_year]])/length(images_year), legend = F, axes = F, box = F, main = i, 
         breaks = c(0,0.0001,0.2,0.4,0.6,0.8,1),
         col = c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5)))
    yearly <- addLayer(yearly, sum(aniBrick_dl[[images_year]])/length(images_year))
}

par(fig = c(0, 1, 0, 1), oma = c(1, 0, 0, 0), mar = c(0, 0, 0, 0), new = TRUE)
plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
legend("bottom", ncol=6, legend=c("Always Land",NA,NA,NA,NA,"Always Water"), xpd = TRUE, horiz = TRUE, inset = c(0, 
                                                                                                                 0), bty="n", fill=c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5)), title="Classification Frequency", cex=1.25)


for(y in 1984:2018){for(m in 5:10){if(length((df %>% filter(year == y) %>% filter(month == m) %>% filter(!row %in% mean_outliers))$row) > 0){seasonal <- addLayer(seasonal, sum(aniBrick_dl[[(df %>% filter(year == y) %>% filter(month == m) %>% filter(!row %in% mean_outliers))$row]])/length((df %>% filter(year == y) %>% filter(month == m) %>% filter(!row %in% mean_outliers))$row))}else{seasonal <- addLayer(seasonal, aniBrick_dl[[1]]*NA)}}}

mains <- c(outer(months, paste(" ", 1984:2018, sep = ""), FUN=paste0))

animate(seasonal, col = colorRampPalette(c("#8ac28c", "#0c28eb"))(20), main = mains)

```