---
title: "Colorado River Analysis"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(raster)
library(robustbase)
library(robust)
library(ggplot2)
```

## Time Series Analysis of Colorado River

Load the set of predictions for the Colorado River four mile square generated using a Random Forest classifier with training on the set of images in "colo_river_training_diff_wl.tif" using a combination of features including MNDWI, NDWI, MIR. 

```{r}
df <- read.csv("colo_river_predictions_lambda0.1.csv")
df <- df %>% mutate(dates = lubridate::ymd(dates)) %>% mutate(jrc = ifelse(dates > lubridate::ymd("2015-12-31"), NA, jrc))
```

## Spatial Analysis

Load images of four mile square

```{r}
aniBrick_cr <- brick("../classification/colo_river_raster_lambda0.1.tif")
aniBrick_cr <- readAll(aniBrick_cr)
```

Recognize anomalous images via high spatial correlation between images with similar number of pixels classified as water, look for high correlation in "extreme weather event" images, if spatial pattern is different it is likely one of those images is anomalous.

```{r}
brick <- aniBrick_cr

yearly <- brick()

for(i in 1984:2018){
    images_year <- (df %>% filter(year == i))$row
    yearly <- addLayer(yearly, sum(brick[[images_year]]))
}

fluctuation_water <- c()
fluctuation_land <- c()
c <- 0
dates <- c()
row <- c()
mean_water <- c()
image_water <- c()
water_brick_cr <- brick()
land_brick_cr <- brick()

for(i in 1984:2018){
  c <- c + 1
  images_year <- (df %>% filter(year == i))$row
  print(i)
  mean <- sum(brick[[images_year]])/length(images_year)
  mean <- mean >= 0.5
  for(j in images_year){
    
    stat_1 <- (brick[[j]] - mean)/(yearly[[c]])
    water <- sum(getValues(stat_1 * (stat_1 > 0)), na.rm = T)
    
    stat_2 <- (1 - brick[[j]] - 1 + mean)/(length(images_year)-yearly[[c]])
    land <- sum(getValues(stat_2 * (stat_2 > 0)), na.rm = T)
    
    water_brick_cr <- addLayer(water_brick_cr, stat_1 * (stat_1 > 0))
    land_brick_cr <- addLayer(land_brick_cr, stat_2 * (stat_2 > 0))
    
    mean_water <- c(mean_water, sum(getValues(mean), na.rm=T))
    image_water <- c(image_water, sum(getValues((brick[[j]])), na.rm=T))
    
    fluctuation_water <- c(fluctuation_water, water)
    fluctuation_land <- c(fluctuation_land, land)
    
    dates <- c(dates, as.character(df[df$row == j,]$dates))
    row <- c(row, j)
  }
}

fluctuation_water_cr <- fluctuation_water
fluctuation_land_cr <- fluctuation_land
mean_water_cr <- mean_water
image_water_cr <- image_water
row_cr <- row
dates_cr <- dates

write.csv(cbind(row_cr, fluctuation_water_cr, fluctuation_land_cr, mean_water_cr, image_water_cr, dates_cr), file = "outlier_analysis_cr.csv")
```

```{r}
t <- read.csv("~/GEE/data/colo_river/outlier_analysis_cr.csv")
brick <- aniBrick_cr
fluctuation_water <- t$fluctuation_water_cr
fluctuation_land <- t$fluctuation_land_cr
mean_water <- t$mean_water_cr
image_water <- t$image_water_cr
row <- t$row_cr
dates <- t$dates_cr

ggplot2::ggplot(data = t %>% mutate(quantile = ntile(fluctuation_water_cr, 10)) %>% group_by(quantile) %>% summarise(mean = mean(fluctuation_water_cr), sd = sd(fluctuation_water_cr)), aes(x = log(mean), y = log(sd))) + ggplot2::geom_point() + ggplot2::geom_smooth(method = "lm") + ggtitle("h_k Box-Cox Plot, Slope: 1.0022")
summary(lm(log(sd) ~ log(mean), data = t %>% mutate(quantile = ntile(fluctuation_water_cr, 10)) %>% group_by(quantile) %>% summarise(mean = mean(fluctuation_water_cr), sd = sd(fluctuation_water_cr))))

######################
resp_water <- fluctuation_water[fluctuation_water>0]
water_mod <- glmrob(resp_water ~ image_water[fluctuation_water>0] +
                        poly(lubridate::yday(dates[fluctuation_water>0]),3),
                    family = Gamma(link="log"))

sum(residuals(water_mod, "deviance")^2)
######################

res <- residuals(water_mod, "deviance") #look only at positive residuals indicating very high water fluctuation
q <- 2

plot(df$pred, x = df$dates, pch = 19)

pal <- colorRampPalette(c("red", "pink"))

points(x = df[(df$row %in% row[fluctuation_water>0][which(res > q)]),]$dates[order(resp_water[which(res > q)], (row[fluctuation_water>0])[which(res > q)], decreasing = T)], y = df[(df$row %in% row[fluctuation_water>0][which(res > q)]),]$pred[order(resp_water[which(res > q)], (row[fluctuation_water>0])[which(res > q)], decreasing = T)], pch = 19, col = pal(length(row[fluctuation_water>0][which(res > q)])))

mean_high_outliers <- df[(df$row %in% row[fluctuation_water>0][which(res > q)]),]$row

mean_land <- ncell(brick[[1]]) - mean_water

###################
resp_land <- fluctuation_land[fluctuation_land>0]
land_mod <- glmrob(resp_land ~ image_water[fluctuation_land>0] +
                       poly(lubridate::yday(dates[fluctuation_land>0]),3) + image_water[fluctuation_land>0]:
                       poly(lubridate::yday(dates[fluctuation_land>0]),1),
                   family = Gamma(link="log"))

sum(residuals(land_mod, "deviance")^2)
###################

res <- residuals(land_mod) #look only at positive residuals indicating very high land fluctuation
q2 <- 2

pal <- colorRampPalette(c("blue", "lightblue"))

points(x = df[(df$row %in% row[fluctuation_land>0][which(res > q2)]),]$dates[order(resp_land[which(res > q2)], (row[fluctuation_land>0])[which(res > q2)], decreasing = T)], y = df[(df$row %in% row[fluctuation_land>0][which(res > q2)]),]$pred[order(resp_land[which(res > q2)], (row[fluctuation_land>0])[which(res > q2)], decreasing = T)], pch = 19, col = pal(length(row[fluctuation_land>0][which(res > q2)])))


mean_low_outliers <- df[(df$row %in% row[fluctuation_land>0][which(res > q2)]),]$row

mean_outliers <- c(mean_high_outliers, mean_low_outliers)

which(df[,'row'] %in% mean_high_outliers)
which(df[,'row'] %in% mean_low_outliers)

```

```{r}

require(gridExtra)
theme_update(plot.title = element_text(hjust = 0.5))
plot1 <- ggplot(df %>% mutate(day = 1) %>% mutate(dates = lubridate::ymd(paste(year,month,day,"-")))) + 
  ggtitle("JRC") +   
  geom_point(aes(x = dates, y = jrc*9/10), size = 0.25) + 
  scale_y_continuous(limits = c(3000,6000))+#max(df$pred*9/10))) +
  ylab(bquote('Water extent (1000 '*m^2*')')) +
  xlab(NULL) +
  scale_x_date(limits = c(lubridate::ymd("2012-01-01"),lubridate::ymd("2015-12-31"))) +
  theme(text = element_text(size=10)) +
  theme(panel.grid.minor = element_blank())
plot2 <- ggplot(df %>% filter(!(row %in% mean_outliers))) + 
  ggtitle(bquote('SWIM '~lambda~'= 0.1')) + 
  geom_point(aes(x = dates, y = pred*9/10), size = 0.25) + 
  scale_y_continuous(limits = c(3000,6000), labels = NULL) +
  ylab(NULL) +
  xlab(NULL) +
  scale_x_date(limits = c(lubridate::ymd("2012-01-01"),lubridate::ymd("2015-12-31"))) +
  theme(text = element_text(size=10)) +
  theme(panel.grid.minor = element_blank())

grid.arrange(plot1, plot2, ncol=2)
timeseries2 <- arrangeGrob(plot1, plot2, ncol=2, widths = c(1.2,1))
ggsave(timeseries2, filename = "colo_river_timeseries_outliers_shown.tiff", dpi = 1200, width = 16.5, height = 8, units = "cm")

```

```{r}

### Three monthly images during drought season

tiff(filename = "colo_river_drought_figure.tiff", width = 8, height = 12, units = "cm", res = 1200)

par(mar = c(0,0,0,0), mfrow = c(3,2), oma = c(7,3,3,0))

y <- 2012; m <- 6; plot(brick('~/GEE/data/colo_river/colo_river_final_extras_rm_jrc.tif')[[(df %>% filter(year == y) %>% filter(month == m) %>% filter(!row %in% mean_outliers))$row[1]]], col = c("white", rgb(92,157,91,maxColorValue=255), "blue"), legend = F, axes = F, box = F)
mtext("JRC", side = 3, cex.main = 1, line = 0.5)
mtext("June 2012", side = 2, cex = 0.75, line = 1)


plot(sum(aniBrick_cr[[(df %>% filter(year == y) %>% filter(month == m) %>% filter(!row %in% mean_outliers))$row]], na.rm=T)/length((df %>% filter(year == y) %>% filter(month == m) %>% filter(!row %in% mean_outliers))$row), col = c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5)), legend = F, axes = F, box = F, breaks = c(0,0.0001,0.2,0.4,0.6,0.8,1));
mtext("SWIM", side = 3, cex.main = 1, line = 0.5)

y <- 2013; m <- 7; plot(brick('~/GEE/data/colo_river/colo_river_final_extras_rm_jrc.tif')[[(df %>% filter(year == y) %>% filter(month == m) %>% filter(!row %in% mean_outliers))$row[1]]]>1, col = c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5)), legend = F, axes = F, box = F)
mtext("July 2013", side = 2, cex = 0.75, line = 1)

plot(sum(aniBrick_cr[[(df %>% filter(year == y) %>% filter(month == m) %>% filter(!row %in% mean_outliers))$row]], na.rm=T)/length((df %>% filter(year == y) %>% filter(month == m) %>% filter(!row %in% mean_outliers))$row), col = c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5)), legend = F, axes = F, box = F, breaks = c(0,0.0001,0.2,0.4,0.6,0.8,1));


y <- 2014; m <- 5; plot(brick('~/GEE/data/colo_river/colo_river_final_extras_rm_jrc.tif')[[(df %>% filter(year == y) %>% filter(month == m) %>% filter(!row %in% mean_outliers))$row[1]]]>1, col = c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5)), legend = F, axes = F, box = F)
mtext("May 2014", side = 2, cex = 0.75, line = 1)

plot(sum(aniBrick_cr[[(df %>% filter(year == y) %>% filter(month == m) %>% filter(!row %in% mean_outliers))$row]], na.rm=T)/length((df %>% filter(year == y) %>% filter(month == m) %>% filter(!row %in% mean_outliers))$row), col = c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5)), legend = F, axes = F, box = F, breaks = c(0,0.0001,0.2,0.4,0.6,0.8,1));


par(fig = c(0, 1, 0, 1), oma = c(1, 0, 1, 0), mar = c(1, 0, 2, 0), new = TRUE)
plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n", maxpixels=1e8)
legend("bottom", ncol=3, legend=c("0%","20%","40%","60%","80%","100%")[c(1,4,2,5,3,6)], xpd = TRUE, inset = c(0, 
    0), bty="n", fill=c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5))[c(1,4,2,5,3,6)], title="Water Frequency", cex=1)

dev.off()

```