---
title: "New Orleans Case Study Analysis"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(raster)
library(robustbase)
library(robust)
library(ggplot2)
```

## Time Series Analysis of New Orleans Spillway

Load the set of predictions for four mile square covering the New Orleans Spillway generated using a Random Forest classifier with training on the set of images 

```{r}
df <- read.csv("new_orleans_predictions_lambda0.3.csv")
df <- df %>% mutate(dates = lubridate::ymd(dates)) %>% mutate(jrc = ifelse(dates > lubridate::ymd("2015-12-31"), NA, jrc))
```

Create a simple time series of the number of pixels classified as water over time. 

```{r}
ggplot(df) + 
  geom_point(aes(x = dates, y = pred*9/10), size = 0.5) + 
  ylab("Water extent (1000 m^2)") +
  xlab("Date") +
  scale_x_date(limits = c(lubridate::ymd("1984-01-01"),lubridate::ymd("2019-12-31"))) +
  theme(text = element_text(size=15))
```

Compare timeseries from JRC and SWIM methods

```{r}
require(gridExtra)
theme_update(plot.title = element_text(hjust = 0.5))
plot1 <- ggplot(df) + 
  ggtitle("JRC") +   
  geom_point(aes(x = dates, y = jrc*9/10), size = 0.5) + 
  scale_y_continuous(limits = c(0,max(df$pred*9/10))) +
  ylab(bquote('Water extent (1000 '*m^2*')')) +
  xlab(NULL) +
  scale_x_date(limits = c(lubridate::ymd("1984-01-01"),lubridate::ymd("2019-12-31"))) +
  theme(text = element_text(size=15))
plot2 <- ggplot(df) + 
  ggtitle("SWIM λ = 0.3") +
  geom_point(aes(x = dates, y = pred*9/10), size = 0.5) + 
  scale_y_continuous(limits = c(0,max(df$pred*9/10)), labels=NULL) +
  ylab(NULL) +
  xlab(NULL) +
  scale_x_date(limits = c(lubridate::ymd("1984-01-01"),lubridate::ymd("2019-12-31"))) +
  theme(text = element_text(size=15))
grid.arrange(plot1, plot2, ncol=2)
timeseries <- arrangeGrob(plot1, plot2, ncol=2)
```

Create a complex graphic with different colors for percent image imputed and a trend connecting only those points with < 1% of the image imputed.

```{r}
rcolors = c('#C8102E', colorRampPalette(c("pink", "blue"))(6))

plot(x = (df %>% filter(cloud < 101))$dates, y = (df %>% filter(cloud < 101))$pred, "p", xlab = "Date", ylab = "Total pixels classified as water", pch = 19, col = rcolors[1], main="Total pixels classified as water after imputation")
points(x = (df %>% filter(cloud < 100))$dates, y = (df %>% filter(cloud < 100))$pred, pch = 19, col = rcolors[2])
points(x = (df %>% filter(cloud < 75))$dates, y = (df %>% filter(cloud < 75))$pred, pch = 19, col = rcolors[3])
points(x = (df %>% filter(cloud < 50))$dates, y = (df %>% filter(cloud < 50))$pred, pch = 19, col = rcolors[4])
points(x = (df %>% filter(cloud < 25))$dates, y = (df %>% filter(cloud < 25))$pred, pch = 19, col = rcolors[5])
points(x = (df %>% filter(cloud < 10))$dates, y = (df %>% filter(cloud < 10))$pred, pch = 19, col = rcolors[6])
points(x = (df %>% filter(cloud < 1))$dates, y = (df %>% filter(cloud < 1))$pred, pch = 19, col = rcolors[7])
lines(x = (df %>% filter(cloud < 1))$dates, y = (df %>% filter(cloud < 1))$pred, "l")
legend(col = rcolors, legend = c("100", "[75, 100)", "[50, 75)", "[25, 50)", "[10, 25)", "[1, 10)", "[0, 1)"), "bottomright", pch = 19, title = "Percent imputed", cex = 1, ncol=1)
```

Histogram of percent image imputed

```{r}
hist(x = df$cloud, xlab = "Percent Messing", main = "Distribution of percentage of image missing")
```

## Spatial Analysis of New Orleans Spillway

Load classified images

```{r}
aniBrick_no <- brick("../classification/new_orleans_raster_lambda0.3.tif")
aniBrick_no <- readAll(aniBrick_no)
```

Plot classified data for all years

```{r}
yearly <- brick()

par(mar = c(0.5,0,1,0), mfrow = c(5,7), oma = c(5,0,0,0))
for(i in 1984:2018){
    images_year <- (df %>% filter(year == i))$row
    par(mar = c(0.5,0,1,0))
    plot(sum(aniBrick_no[[images_year]])/length(images_year), legend = F, axes = F, box = F, main = i, 
         breaks = c(0,0.0001,0.2,0.4,0.6,0.8,1),
         col = c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5)))
    yearly <- addLayer(yearly, sum(aniBrick_no[[images_year]])/length(images_year))
}

par(fig = c(0, 1, 0, 1), oma = c(1, 0, 0, 0), mar = c(0, 0, 0, 0), new = TRUE)
plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
legend("bottom", ncol=6, legend=c("Always Land",NA,NA,NA,NA,"Always Water"), xpd = TRUE, horiz = TRUE, inset = c(0, 
    0), bty="n", fill=c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5)), title="Classification Frequency", cex=1.25)
```

Produce animation

```{r}
par(mfrow = c(1,1), mar = c(1,1,1,1))
animate(yearly, main = c(1984:2018), n = 1, pause = 1,
        breaks = c(0,0.0001,0.2,0.4,0.6,0.8,1),
        lab.breaks = c("Always\nland",NA,NA,NA,NA,NA,"Always\nwater"),
        col = c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5)))
```

Recognize anomalous images via high spatial correlation between images with similar number of pixels classified as water, look for high correlation in "extreme weather event" images, if spatial pattern is different it is likely one of those images is anomalous.

```{r}
brick <- aniBrick_no

yearly <- brick()

for(i in 1984:2018){
    images_year <- (df %>% filter(year == i))$row
    yearly <- addLayer(yearly, sum(brick[[images_year]]))
}

fluctuation_water <- c()
fluctuation_land <- c()
c <- 0
dates <- c()
row <- c()
mean_water <- c()
image_water <- c()
water_brick_no <- brick()
land_brick_no <- brick()

for(i in 1984:2018){
  c <- c + 1
  images_year <- (df %>% filter(year == i))$row
  print(i)
  mean <- sum(brick[[images_year]])/length(images_year)
  mean <- mean >= 0.5
  for(j in images_year){
    
    stat_1 <- (brick[[j]] - mean)/(yearly[[c]])
    water <- sum(getValues(stat_1 * (stat_1 > 0)), na.rm = T)
    
    stat_2 <- (1 - brick[[j]] - 1 + mean)/(length(images_year)-yearly[[c]])
    land <- sum(getValues(stat_2 * (stat_2 > 0)), na.rm = T)
    
    water_brick_no <- addLayer(water_brick_no, stat_1 / stat_1)
    land_brick_no <- addLayer(land_brick_no, stat_2 / stat_2)
    
    mean_water <- c(mean_water, sum(getValues(mean), na.rm=T))
    image_water <- c(image_water, sum(getValues((brick[[j]])), na.rm=T))
    
    fluctuation_water <- c(fluctuation_water, water)
    fluctuation_land <- c(fluctuation_land, land)
    
    dates <- c(dates, as.character(df[df$row == j,]$dates))
    row <- c(row, j)
  }
}

fluctuation_water_no <- fluctuation_water
fluctuation_land_no <- fluctuation_land
mean_water_no <- mean_water
image_water_no <- image_water
row_no <- row
dates_no <- dates

write.csv(cbind(row_no, fluctuation_water_no, fluctuation_land_no, mean_water_no, image_water_no, dates_no), file = "outlier_analysis_no.csv")

```

```{r}
t <- read.csv("outlier_analysis_no.csv")
brick <- aniBrick_no
fluctuation_water <- t$fluctuation_water_no
fluctuation_land <- t$fluctuation_land_no
mean_water <- t$mean_water_no
image_water <- t$image_water_no
row <- t$row_no
dates <- t$dates_no

######################
resp_water <- fluctuation_water[fluctuation_water>0]
water_mod <- glmrob(resp_water ~ image_water[fluctuation_water>0] +
                        poly(lubridate::yday(dates[fluctuation_water>0]), 3),
                    family = Gamma(link="log"))

sum(residuals(water_mod, "deviance")^2)
######################

res <- residuals(water_mod, "deviance")
q <- 2

plot(df$pred, x = df$dates, pch = 19)

pal <- colorRampPalette(c("red", "pink"))

points(x = df[(df$row %in% row[fluctuation_water>0][which(res > q)]),]$dates[order(resp_water[which(res > q)], (row[fluctuation_water>0])[which(res > q)], decreasing = T)], y = df[(df$row %in% row[fluctuation_water>0][which(res > q)]),]$pred[order(resp_water[which(res > q)], (row[fluctuation_water>0])[which(res > q)], decreasing = T)], pch = 19, col = pal(length(row[fluctuation_water>0][which(res > q)])))

mean_high_outliers <- df[(df$row %in% row[fluctuation_water>0][which(res > q)]),]$row

mean_land <- ncell(brick[[1]]) - mean_water

###################
resp_land <- fluctuation_land[fluctuation_land>0]
land_mod <- glmrob(resp_land ~ image_water[fluctuation_land>0] +
                       poly(lubridate::yday(dates[fluctuation_land>0]),3),
                   family = Gamma(link="log"))

sum(residuals(land_mod, "deviance")^2)
###################

res <- residuals(land_mod)
q2 <- 2

pal <- colorRampPalette(c("blue", "lightblue"))

points(x = df[(df$row %in% row[fluctuation_land>0][which(res > q2)]),]$dates[order(resp_land[which(res > q2)], (row[fluctuation_land>0])[which(res > q2)], decreasing = T)], y = df[(df$row %in% row[fluctuation_land>0][which(res > q2)]),]$pred[order(resp_land[which(res > q2)], (row[fluctuation_land>0])[which(res > q2)], decreasing = T)], pch = 19, col = pal(length(row[fluctuation_land>0][which(res > q2)])))


mean_low_outliers <- df[(df$row %in% row[fluctuation_land>0][which(res > q2)]),]$row

mean_outliers <- c(mean_high_outliers, mean_low_outliers)

which(df[,'row'] %in% mean_high_outliers)
which(df[,'row'] %in% mean_low_outliers)
```

```{r}
require(gridExtra)
theme_update(plot.title = element_text(hjust = 0.5))
plot1 <- ggplot(df) + 
    ggtitle("JRC") +   
    geom_point(aes(x = dates, y = jrc*9/10), size = 0.25) + 
    scale_y_continuous(limits = c(0,max(df$pred*9/10))) +
    ylab(bquote('Water extent (1000 '*m^2*')')) +
    xlab(NULL) +
    scale_x_date(limits = c(lubridate::ymd("1985-01-01"),lubridate::ymd("2014-12-31"))) +
    theme(text = element_text(size=10))
plot2 <- ggplot(df %>% filter(!(row %in% mean_outliers))) + 
    ggtitle(bquote('SWIM '~lambda~'= 0.3')) + 
    geom_point(aes(x = dates, y = pred*9/10), size = 0.25) + 
    scale_y_continuous(limits = c(0,max(df$pred*9/10)), labels=NULL) +
    ylab(NULL) +
    xlab(NULL) +
    scale_x_date(limits = c(lubridate::ymd("1985-01-01"),lubridate::ymd("2014-12-31"))) +
    theme(text = element_text(size=10))
grid.arrange(plot1, plot2, ncol=2)

timeseries2 <- arrangeGrob(plot1, plot2, ncol=2, widths = c(1.2,1))
```

```{r}
require(gridExtra)
theme_update(plot.title = element_text(hjust = 0.5))
plot1 <- ggplot(df) + 
  ggtitle("JRC") +   
  geom_point(aes(x = dates, y = jrc*9/10), size = 0.5) + 
  scale_y_continuous(limits = c(0,max(df$pred*9/10))) +
  ylab(bquote('Water extent (1000 '*m^2*')')) +
  xlab(NULL) +
  scale_x_date(limits = c(lubridate::ymd("1984-01-01"),lubridate::ymd("2019-12-31"))) +
  theme(text = element_text(size=15))
plot2 <- ggplot(df) + 
  ggtitle(bquote('CSSM, '~lambda~'= 0.3')) + 
  geom_point(aes(x = dates, y = pred*9/10), size = 0.5) + 
  scale_y_continuous(limits = c(0,max(df$pred*9/10)), labels=NULL) +
  ylab(NULL) +
  xlab(NULL) +
  scale_x_date(limits = c(lubridate::ymd("1984-01-01"),lubridate::ymd("2019-12-31"))) +
  theme(text = element_text(size=15))
plot3 <- ggplot(df %>% filter(!(row %in% mean_outliers))) + 
  ggtitle("CSSM identify outliers") +
  geom_point(aes(x = dates, y = pred*9/10), size = 0.5) + 
  geom_point(data = df %>% filter((row %in% mean_outliers)), aes(x = dates, y = pred*9/10), colour="red", size=0.5, shape=4) +
  scale_y_continuous(limits = c(0,max(df$pred*9/10))) +
  ylab(bquote('Water extent (1000 '*m^2*')')) +
  xlab(NULL) +
  scale_x_date(limits = c(lubridate::ymd("1984-01-01"),lubridate::ymd("2019-12-31"))) +
  theme(text = element_text(size=15))
plot4 <- ggplot(df %>% filter(!(row %in% mean_outliers))) + 
  ggtitle("CSSM outliers removed") +
  geom_point(aes(x = dates, y = pred*9/10), size = 0.5) + 
  scale_y_continuous(limits = c(0,max(df$pred*9/10)), labels=NULL) +
  ylab(NULL) +
  xlab(NULL) +
  scale_x_date(limits = c(lubridate::ymd("1984-01-01"),lubridate::ymd("2019-12-31"))) +
  theme(text = element_text(size=15))
grid.arrange(plot1, plot2, plot3, plot4, ncol=2, widths = c(1.2,1))
timeseries3 <- arrangeGrob(plot1, plot2, plot3, plot4, ncol=2, widths = c(1.2,1))
```

```{r}
png(filename = "new_orleans_spatial_outliers_rm.png", width = 8.5, height = 8.5, units = "in", res = 1200)

yearly <- brick()

par(mar = c(0.5,0,1,0), mfrow = c(5,7), oma = c(5,0,0,0))
for(i in 1984:2018){
    images_year <- (df %>% filter(year == i) %>% filter(!(row %in% mean_outliers)))$row
    par(mar = c(0.5,0,1,0))
    plot(sum(aniBrick_no[[images_year]])/length(images_year), legend = F, axes = F, box = F, main = i, 
         breaks = c(0,0.0001,0.2,0.4,0.6,0.8,1),
         col = c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5)))
    yearly <- addLayer(yearly, sum(aniBrick_no[[images_year]])/length(images_year))
}

par(fig = c(0, 1, 0, 1), oma = c(1, 0, 0, 0), mar = c(0, 0, 0, 0), new = TRUE)
plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
legend("bottom", ncol=6, legend=c("0%","20%","40%","60%","80%","100%"), xpd = TRUE, horiz = TRUE, inset = c(0, 
    0), bty="n", fill=c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5)), title="Water Frequency", cex=1.25)

dev.off()
```

```{r}

png(filename = "new_orleans_flood_figure.png", width = 5, height = 9, units = "in", res = 1200)

par(mar = c(0,0,0,0), mfrow = c(3,2), oma = c(10,5,0,0))

y <- 1994; plot(sum(brick('~/GEE/data/new_orleans/new_orleans_final_extras_rm_jrc.tif')[[(df %>% filter(year == y) %>% filter(!row %in% mean_outliers))$row]]>1)/length((df %>% filter(year == y) %>% filter(!row %in% mean_outliers))$row), col = c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5)), legend = F, axes = F, box = F, breaks = c(0,0.0001,0.2,0.4,0.6,0.8,1))
title("JRC", cex.main = 2, line = -2)
mtext("1994", side = 2, cex = 1.5, line = 1)

plot(sum(brick[[(df %>% filter(year == y) %>% filter(!row %in% mean_outliers))$row]])/length((df %>% filter(year == y) %>% filter(!row %in% mean_outliers))$row), col = c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5)), legend = F, axes = F, box = F, breaks = c(0,0.0001,0.2,0.4,0.6,0.8,1))
title("SWIM", cex.main = 2, line = -2)

y <- 1997; plot(sum(brick('~/GEE/data/new_orleans/new_orleans_final_extras_rm_jrc.tif')[[(df %>% filter(year == y) %>% filter(!row %in% mean_outliers))$row]]>1)/length((df %>% filter(year == y) %>% filter(!row %in% mean_outliers))$row), col = c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5)), legend = F, axes = F, box = F, breaks = c(0,0.0001,0.2,0.4,0.6,0.8,1))
mtext("1997", side = 2, cex = 1.5, line = 1)

plot(sum(brick[[(df %>% filter(year == y) %>% filter(!row %in% mean_outliers))$row]])/length((df %>% filter(year == y) %>% filter(!row %in% mean_outliers))$row), col = c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5)), legend = F, axes = F, box = F, breaks = c(0,0.0001,0.2,0.4,0.6,0.8,1))


y <- 2011; plot(sum(brick('~/GEE/data/new_orleans/new_orleans_final_extras_rm_jrc.tif')[[(df %>% filter(year == y) %>% filter(!row %in% mean_outliers))$row]]>1)/length((df %>% filter(year == y) %>% filter(!row %in% mean_outliers))$row), col = c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5)), legend = F, axes = F, box = F, breaks = c(0,0.0001,0.2,0.4,0.6,0.8,1))
mtext("2011", side = 2, cex = 1.5, line = 1)

plot(sum(brick[[(df %>% filter(year == y) %>% filter(!row %in% mean_outliers))$row]])/length((df %>% filter(year == y) %>% filter(!row %in% mean_outliers))$row), col = c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5)), legend = F, axes = F, box = F, breaks = c(0,0.0001,0.2,0.4,0.6,0.8,1))


par(fig = c(0, 1, 0, 1), oma = c(1, 0, 1, 0), mar = c(1, 0, 2, 0), new = TRUE)
plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n", maxpixels=1e8)
legend("bottom", ncol=3, legend=c("0%","20%","40%","60%","80%","100%")[c(1,4,2,5,3,6)], xpd = TRUE, inset = c(0, 
    0), bty="n", fill=c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5))[c(1,4,2,5,3,6)], title="Water Frequency", cex=2)

dev.off()

```

```{r}

tiff(filename = "new_orleans_flood_figure.tiff", width = 8, height = 15, units = "cm", res = 1200)

par(mar = c(0,0,0,0), mfrow = c(3,2), oma = c(10,5,0,0))

y <- 1994; plot(sum(brick('~/GEE/data/new_orleans/new_orleans_final_extras_rm_jrc.tif')[[(df %>% filter(year == y) %>% filter(!row %in% mean_outliers))$row]]>1)/length((df %>% filter(year == y) %>% filter(!row %in% mean_outliers))$row), col = c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5)), legend = F, axes = F, box = F, breaks = c(0,0.0001,0.2,0.4,0.6,0.8,1))
title("JRC", cex.main = 2, line = -2)
mtext("1994", side = 2, cex = 1.5, line = 1)

plot(sum(brick[[(df %>% filter(year == y) %>% filter(!row %in% mean_outliers))$row]])/length((df %>% filter(year == y) %>% filter(!row %in% mean_outliers))$row), col = c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5)), legend = F, axes = F, box = F, breaks = c(0,0.0001,0.2,0.4,0.6,0.8,1))
title("SWIM", cex.main = 2, line = -2)

y <- 1997; plot(sum(brick('~/GEE/data/new_orleans/new_orleans_final_extras_rm_jrc.tif')[[(df %>% filter(year == y) %>% filter(!row %in% mean_outliers))$row]]>1)/length((df %>% filter(year == y) %>% filter(!row %in% mean_outliers))$row), col = c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5)), legend = F, axes = F, box = F, breaks = c(0,0.0001,0.2,0.4,0.6,0.8,1))
mtext("1997", side = 2, cex = 1.5, line = 1)

plot(sum(brick[[(df %>% filter(year == y) %>% filter(!row %in% mean_outliers))$row]])/length((df %>% filter(year == y) %>% filter(!row %in% mean_outliers))$row), col = c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5)), legend = F, axes = F, box = F, breaks = c(0,0.0001,0.2,0.4,0.6,0.8,1))


y <- 2011; plot(sum(brick('~/GEE/data/new_orleans/new_orleans_final_extras_rm_jrc.tif')[[(df %>% filter(year == y) %>% filter(!row %in% mean_outliers))$row]]>1)/length((df %>% filter(year == y) %>% filter(!row %in% mean_outliers))$row), col = c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5)), legend = F, axes = F, box = F, breaks = c(0,0.0001,0.2,0.4,0.6,0.8,1))
mtext("2011", side = 2, cex = 1.5, line = 1)

plot(sum(brick[[(df %>% filter(year == y) %>% filter(!row %in% mean_outliers))$row]])/length((df %>% filter(year == y) %>% filter(!row %in% mean_outliers))$row), col = c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5)), legend = F, axes = F, box = F, breaks = c(0,0.0001,0.2,0.4,0.6,0.8,1))


par(fig = c(0, 1, 0, 1), oma = c(1, 0, 1, 0), mar = c(1, 0, 2, 0), new = TRUE)
plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n", maxpixels=1e8)
legend("bottom", ncol=3, legend=c("0%","20%","40%","60%","80%","100%")[c(1,4,2,5,3,6)], xpd = TRUE, inset = c(0, 
    0), bty="n", fill=c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5))[c(1,4,2,5,3,6)], title="Water Frequency", cex=1.5)

dev.off()

```

```{r}

tiff(filename = "new_orleans_flood_figure.tiff", width = 8, height = 8, units = "cm", res=1200)

par(mar = c(0,0,0,0), mfrow = c(1,2), oma = c(10,5,10,0))

y <- 1997; plot(sum(brick('~/GEE/data/new_orleans/new_orleans_final_extras_rm_jrc.tif')[[(df %>% filter(year == y) %>% filter(!row %in% mean_outliers))$row]]>1)/length((df %>% filter(year == y) %>% filter(!row %in% mean_outliers))$row), col = c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5)), legend = F, axes = F, box = F, breaks = c(0,0.0001,0.2,0.4,0.6,0.8,1))
#mtext("1997", side = 2, cex = 1, line = 1)
title("JRC", cex.main = 2, line = -2)

plot(sum(brick[[(df %>% filter(year == y) %>% filter(!row %in% mean_outliers))$row]])/length((df %>% filter(year == y) %>% filter(!row %in% mean_outliers))$row), col = c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5)), legend = F, axes = F, box = F, breaks = c(0,0.0001,0.2,0.4,0.6,0.8,1))
title("SWIM", cex.main = 2, line = -2)

par(fig = c(0, 1, 0, 1), oma = c(1, 0, 1, 0), mar = c(1, 0, 2, 0), new = TRUE)
plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n", maxpixels=1e8)
legend("bottom", ncol=3, legend=c("0%","20%","40%","60%","80%","100%")[c(1,4,2,5,3,6)], xpd = TRUE, inset = c(0, 
    0), bty="n", fill=c(rgb(92,157,91,maxColorValue=255), colorRampPalette(c("lightblue", "blue"))(5))[c(1,4,2,5,3,6)], title="Water Frequency", cex=1)

dev.off()

```

```{r}
seasonal <- brick()

for(y in 1984:2018){for(m in 5:10){if(length((df %>% filter(year == y) %>% filter(month == m) %>% filter(!row %in% mean_outliers))$row) > 0){seasonal <- addLayer(seasonal, sum(aniBrick_no[[(df %>% filter(year == y) %>% filter(month == m) %>% filter(!row %in% mean_outliers))$row]])/length((df %>% filter(year == y) %>% filter(month == m) %>% filter(!row %in% mean_outliers))$row))}else{seasonal <- addLayer(seasonal, aniBrick_no[[1]]*NA)}}}

mains <- c(outer(months, paste(" ", 1984:2018, sep = ""), FUN=paste0))

animate(seasonal, col = colorRampPalette(c("#8ac28c", "#0c28eb"))(20), main = mains)

```